### 2018-05-18 paper/intro.tex
Do some copyediting and typesetting

\section{Introduction}

In the two decades since they were introduced by \citet{wadler89}, \emph{type classes} have proven a versatile apparatus for structuring programs. Consider a class intended to characterize types $\tVar$ for which a monoid can be defined:
\begin{flalign*}
  & \eClass{\tMonoid{\tVar}} & \\
  & \quad \tAnno{\eIdentity}{\tVar} & \\
  & \quad \tAnno{\eCombine}{\tArrow{\tVar}{\tArrow{\tVar}{\tVar}}} &
\end{flalign*}
Programmers can use this class to define polymorphic functions over all monoids:
\begin{flalign*}
  & \tAnno{\eSquare}{\tMonoid{\tVar} \Rightarrow \tArrow{\tVar}{\tVar}} & \\
  & \eApp{\eSquare}{\eX} = \eApp{\eApp{\eCombine}{\eX}}{\eX} &
\end{flalign*}
Finally, an \emph{instance} of the class can be provided for a specific type:
\begin{flalign*}
  & \eInstance{\tMonoid{\integer}} & \\
  & \quad \eIdentity = \numberName{1} & \\
  & \quad \eCombine = \eMultiply &
\end{flalign*}
where $\tAnno{\numberName{1}}{\integer}$ and $\tAnno{\eMultiply}{\tArrow{\integer}{\tArrow{\integer}{\integer}}}$. The instance is a witness to justify the well-definedness of expressions like $\eApp{\eSquare}{\numberName{3}}$.

More recently, algebraic effects \citep{plotkin03} and handlers \citep{plotkin09} have become an attractive alternative to monads for modeling computational effects due to their compositionality.